### 2015-02-14 paper.tex
separated content from paper

\documentclass[11pt]{article} % For LaTeX2e
\usepackage{rldmsubmit,palatino}
\usepackage{graphicx}
\usepackage{caption}
\usepackage{subcaption}

\title{Modular Inverse Reinforcement Learning on Human Motion}

\author{
Shun Zhang\\
Department of Computer Science\\
University of Texas at Austin\\
Austin, TX 78712 \\
\texttt{menie482@cs.utexas.edu} \\
\And
Matthew Tong \\
Center for Perceptual Systems\\
University of Texas at Austin\\
Austin, TX 78712 \\
\texttt{mhtong@gmail.com} \\
\AND
Mary Hayhoe \\
Center for Perceptual Systems\\
University of Texas at Austin\\
Austin, TX 78712 \\
\texttt{hayhoe@utexas.edu} \\
\And
Dana Ballard \\
Department of Computer Science\\
University of Texas at Austin\\
Austin, TX 78712 \\
\texttt{dana@cs.utexas.edu} \\
\\
}

% The \author macro works with any number of authors. There are two commands
% used to separate the names and addresses of multiple authors: \And and \AND.
%
% Using \And between authors leaves it to \LaTeX{} to determine where to break
% the lines. Using \AND forces a linebreak at that point. So, if \LaTeX{}
% puts 3 of 4 authors names on the first line, and the last on the second
% line, try using \AND instead of \And before the third author name.

\newcommand{\fix}{\marginpar{FIX}}
\newcommand{\new}{\marginpar{NEW}}

\newcommand{\commentd}[1]{{\bf **Dana: #1**}}

\begin{document}

\maketitle

%The abstract should be a maximum of 2000 characters of text, including
%spaces (no figure is allowed).
\begin{abstract}
Human learning is modeled as reinforcement learning in the literature. However,
human is able to learn complex tasks efficiently, while reinforcement
learning is known for suffering from the curse of dimensionality. We observe that
tasks that human handles can usually be decomposed into smaller tasks, and human can
learn small tasks quickly, and then learn how to integrate them to solve the
original task. In this paper, we find that modular reinforcement learning can be
used to model human behavior. We report experiments of using modular inverse
reinforcement learning to find the way that human combines sub-tasks, given
samples of real human data. Our simulation demonstrates that human behavior can
be well explained in this way.
\end{abstract}

\keywords{
Reinforcement learning, human motion, inverse learning
}

\startmain % to start the main 1-4 pages of the submission.

\include{content}

\bibliographystyle{plain}
\bibliography{paper}

\end{document}