### 2015-03-28 5-conclusion.tex

\section{Conclusions}
\label{sec:conclude}

We analyzed human behavior using inverse modular reinforcement learning. The
experimental results show that modular reinforcement learning can explain human
behavior well, even though the performance of the agent is currently inferior to
human subjects'.

Note that a weighted sum of Q functions is just one way to combine multiple
sub-MDPs. Other ways are possible, including, for example, scheduling between
different modules, with only one active at one time. This is also called {\em
skill} in the literature \cite{konidaris2009skill}. However, we adopt the
weighted sum approach because this is more reasonable for human behavior. When a
human tries to collect targets while avoiding obstacles, these two modules are
expected to be both active. A scheduling approach may yield frequent oscillation
between these two modules. Note also that we assume independence between
modules. However, correlation between modules doesn't impair our analysis in
this paper. In Figure~\ref{fig:heatmap}, we can tell that the target module and
obstacle module tend to be negatively correlated from the shape of the white
zones. Lastly, weights may be dynamic and different from state to state.
However, with such an assumption we need to learn a mapping from state to
weights. In this case, the curse of dimensionality still exists, and inverse
learning would be difficult.