### 2015-03-29 3-irl.tex
minor changes in irl

\section{Modular Inverse Reinforcement Learning}
\label{sec:rl}

% TODO expand this
In Reinforcement Learning, when given a state and action pair, $s$ and $a$, the
function $Q(s, a)$ evaluates the utility of taking action $a$ from state $s$.
The {\em policy} of a state is the action with the maximum Q value \cite{rl}. So
how do we determine the global policy from the modules, or sub-tasks?

In the literature, modular formation is proposed for understanding human
visuomotor behavior \cite{sprague2007modeling}. A combination of various basic
modules can generate complex behavior to match human behavior.  The combination
can be either of Q functions of modules \cite{rothkopf2013modular} or of rewards
of modules \cite{Rothkopf12Infer}.
Other work has been done to integrate the decomposed value functions of
the sub-tasks \cite{koller1999computing}. The sub-tasks can also propose their
own policies, while the global policy is a weighted sum of those sub-task
policies \cite{thomas2012motor}.

We now consider how the
modules can be combined. Here we assume that the global Q function is a weighted
sum of all $Q_i$ \cite{sprague2007modeling, rothkopf2013modular}, where $Q_i$ is
the Q function for i-th module.
$$Q(s, a) = \sum_i w_i Q_i (s_i, a)$$
where $w_i$ is the weight of the i-th sub-MDP. $w_i \geq 0, \sum_i w_i = 1$.
$s_i$ denotes the decomposition of $s$ in the i-th module.

Different weights can yield different performance. Let $w$ be the vector of
$(w_1, w_2, w_3)$, where $w_1, w_2, w_3$ are weights for the sub-tasks of target et al 
collection, obstacle avoidance, and path following, respectively. An agent with
$w = (1, 0, 0)$ only collects targets, and one with $w = (0, 0.5, 0.5)$ may
avoid the obstacles and follow the path.

% TODO mention discounters here
To obtain the weights given the samples of observed human behavior, we use inverse
reinforcement learning. In reinforcement learning, we derive 
optimal policies given an MDP. In inverse reinforcement learning, however, aims to
find out the underlying MDP by observing policies. A common
way is to use a maximum likelihood method to recover the transition function and
the reward function \cite{ng2000algorithms}, but this approach is very expensive
as some kind of learning is used in the innermost search loop. The modular
approach allows a much more economical approach \cite{rothkopf2013modular},
since we module weights are the only free parameters. The transition function is
known, and the reward function is trivially the weighted sum of that of modules.
More precisely, we use the modular inverse reinforcement learning method in
\cite{rothkopf2013modular} to maximize the function below.
% TODO more details
\begin{equation}
\label{eq:irl}
\max_w \prod_t \frac{e^{\eta Q(s^{(t)}, a^{(t)})}}{\sum_b e^{\eta Q(s^{(t)}, b)}}
\end{equation}
where $s^{(t)}$ is the state at time $t$, and $a^{(t)}$ is the action at time
$t$, which are both from samples. $Q(s, a) = \sum_i w_i Q_i(s_i, a)$, as defined
before. $\eta$ is a hyperparameter that determines the consistency of human's
behavior. The larger $\eta$ is, the algorithm is more likely to overfit the data.

The intuition of Equation~\ref{eq:irl} is that if an action is observed from the
sample, then the Q value of taking that action should be larger compared to Q
values of taking other actions.